Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-05-10 Grassmannian_cells.tex
Edited Section 2
Try to finish it next week, need to proofread it before

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Torus Action on Quiver Grassmannians}
\noindent Let $Q$ be an acylic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:i\to j$. Moreover, let $W_{Q}$ be the free (non-abelian) group generated by $Q_1$.  We denote by $\Rep(Q)$ the category of $\C$-representations of $Q$.
\begin{definition}
The universal covering quiver $\tilde Q$ of $Q$ is given by the vertices $\tilde Q_0=Q_0\times W_{Q}$ and the arrow set $\tilde Q_1=Q_1\times W_{Q}$ where $(\alpha,w):(i,w)\to (j,w\alpha)$ for every $\alpha:i\to j$.

We say that a representation $X\in \Rep(Q)$ can be lifted (to $\tilde Q$) if there exists a representation $\tilde X\in\Rep(\tilde Q)$ such that $F_Q\tilde X=X$ where $F_Q:\Rep(\tilde Q)\to\Rep(Q)$ is the natural functor.
\end{definition}
\begin{lemma}
  Every preprojective (resp. preinjective) representation of $Q$ can be lifted to a representation of $\tilde Q$.
\end{lemma}
\begin{proof}This statement is clear for the simple representations $S_q$, $q\in Q_0$. Now every preprojective representation $X$ can be obtained when applying a sequence of BGP-reflections \cite{bgp} to a simple representation $S_{q'}$ of a quiver $Q'$ whose underlying graph is the one of $Q$. Applying BGP-reflections to a source $q$ of $Q$ corresponds to applying BGP-reflections to all vertices $(q,w)$ of $\tilde Q$. This leads the claim. The statement for preinjective representations follows in the same way.
\end{proof}
We choose a map $d:\tilde Q_0\to\ZZ$ and fix a representation $X\in\Rep(Q)$. In any case, we can consider the decomposition $X_q=\bigoplus_{w\in W_Q} X_{(q,w)}$. We define a torus action on each $X_{(q,w)}$ via $t.x_{(q,w)}=t^{d(q,w)}x_{(q,w)}$ which can be extended linearly to each $X_q$. For a fixed a subspace $U_q$, we can define the subspace $t.U_q$. In general, this torus action induces no torus action on the Quiver Grassmannians $\Gr_{\bfe}(X)$ as $t.U=(t.U_q)_{q\in Q_0}$ is no subrepresentation of $X$ for every $U\in \Gr_{\bfe}(X)$. Actually, for this the action has to satisfy $X_{\alpha}(t.U_i)\in t.U_j$ for every $\alpha:i\to j$ and every $x\in X_i$. 
\begin{lemma}Fix an integer $c_\alpha\in\ZZ$ for every $\alpha\in Q_1$.
If $X\in\Rep(Q)$ can be lifted, $d:\tilde Q_0\to\ZZ$ induces a torus action on $\Gr_\bfe(X)$ if we have $d(j,w\alpha)-d(i,w)=c_\alpha$ for all $\alpha:i\to j$ and $w\in W_{Q}$.
\end{lemma}
\begin{proof} Since $X$ can be lifted, we can write $X_\alpha:X_i\to X_j$ as block matrix consisting of linear maps $X_{(\alpha,w)}:X_{(i,w)}\to X_{(j,w\alpha)}$. Then the condition $X_{\alpha}(t.U_i)\in t.U_j$ translates into ...to be continued.
\end{proof}

\begin{lemma}
  There exists $d:\tilde Q_0\to\ZZ$ such that $d(q,w)\ne d(q',w')$ for all $q,q',w,w'$ with $\dim X_{q,w}\ne0$ and $\dim X_{q',w'}\ne0$.
\end{lemma}

\begin{theorem}
  $Gr^Q_\bfe(X)^T\cong\bigsqcup_{\tilde\bfe of type \bfe} Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$
\end{theorem}

\begin{corollary}
  affine bundles over $Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$ if $Gr^Q_\bfe(X)$ is smooth
  \[\{U\in Gr^Q_\bfe(X):\lim_{t\to0} t\cdot U\in Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)\}\]
\end{corollary}

@@

\bibitem{sch} Schofield, A.: General representations of quivers. Proceedings of the London Mathematical Society (3) \textbf{65}(1), 46-64 (1992).

\end{thebibliography}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Torus Action on Quiver Grassmannians}
\noindent Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:i\to j$. We denote by $\Rep(Q)$ the category of $\C$-representations of $Q$. Moreover, let $A_Q\cong \ZZ^{Q_1}$ be the free abelian group and $W_{Q}$ be the free non-abelian group generated by $Q_1$.  
\begin{definition}
The universal abelian covering quiver $\hat Q$ of $Q$ is given by the vertices $\hat Q_0=Q_0\times A_{Q}$ and the arrow set $\hat Q_1=Q_1\times A_{Q}$ where $(\alpha,\chi):(i,\chi)\to (j,\chi+e_\alpha)$ for every $\alpha:i\to j$.

The universal covering quiver $\tilde Q$ of $Q$ is given by the vertices $\tilde Q_0=Q_0\times W_{Q}$ and the arrow set $\tilde Q_1=Q_1\times W_{Q}$ where $(\alpha,w):(i,w)\to (j,w\alpha)$ for every $\alpha:i\to j$.

We say that a representation $X\in \Rep(Q)$ can be lifted to $\hat Q$ (resp. $\tilde Q$) if there exists a representation $\hat X\in\Rep(\hat Q)$ (resp. $\tilde X\in\Rep(\tilde Q)$) such that $F_Q\hat X=X$ (resp. $G_Q \tilde X=X$) where $F_Q:\Rep(\hat Q)\to\Rep(Q)$ (resp. $G_Q:\Rep(\tilde Q)\to\Rep(Q)$) is the natural functor.
\end{definition}
Note that we also have a natural functor $H_Q:\Rep(\tilde Q)\to \Rep(\hat Q)$.
Moreover, note that for every connected component of the universal covering quiver of the universal abelian covering quiver is isomorphic to a connected component of the universal covering quiver of the original quiver. The functor $F_Q$ induces a map $F_Q:\ZZ^{\hat Q_1}\to \ZZ^{Q_1}$. We say that a dimension vector $\hat \bfe$ of $\hat Q$ is compatible with $\bfe$ if $F_Q(\hat\bfe)=\bfe$. Moreover, the group $A_Q$ acts on $\hat Q$ via translation inducing an action on $\Rep(\hat Q)$ and on $\ZZ^{\hat Q_1}$. We say two representation (resp. dimension vectors) are equivalent in this case. The analogous observation can also be made for $\tilde Q$.

\begin{lemma}
  Every preprojective (resp. preinjective) representation of $Q$ can be lifted to  $\hat Q$ and $\tilde Q$.
\end{lemma}
\begin{proof}This statement is clear for the simple representations $S_q$, $q\in Q_0$. Now every preprojective representation $X$ can be obtained when applying a sequence of BGP-reflections \cite{bgp} to a simple representation $S_{q'}$ of a quiver $Q'$ whose underlying graph is the same as the one of $Q$. Applying BGP-reflections to a source $q$ of $Q$ corresponds to applying BGP-reflections to all vertices $(q,\chi)$ of $\hat Q$ (resp. $(q,w)$ of $\tilde Q$). This leads the claim. The statement for preinjective representations follows in the same way.
\end{proof}

We choose a map $d:\hat Q_0\to\ZZ$ and fix a representation $X\in\Rep(Q)$ which can be lifted to $\hat Q$. We can consider the decomposition $X_q=\bigoplus_{(q,\chi)\in A_Q} X_{(q,\chi)}$. We define a torus action on each $X_{(q,\chi)}$ via $t.x_{(q,\chi)}=t^{d(q,\chi)}x_{(q,\chi)}$ which can be extended linearly to each $X_q$. For a fixed a subspace $U_q$, we can define the subspace $t.U_q$. In general, this torus action induces no torus action on the Quiver Grassmannians $\Gr_{\bfe}(X)$ as $t.U=(t.U_q)_{q\in Q_0}$ is no subrepresentation of $X$ for every $U\in \Gr_{\bfe}(X)$. Actually, for this the action has to satisfy $X_{\alpha}(t.U_i)\in t.U_j$ for every $\alpha:i\to j$ and every $x\in X_i$. 
\begin{lemma}Fix an integer $c_\alpha\in\ZZ$ for every $\alpha\in Q_1$.
If $X\in\Rep(Q)$ can be lifted, $d:\hat Q_0\to\ZZ$ induces a torus action on $\Gr_\bfe(X)$ if we have $d(j,\chi+e_\alpha)-d(i,\chi)=c_\alpha$ for all $\alpha:i\to j$ and $w\in W_{Q}$.
\end{lemma}
\begin{proof} Let $U\in \Gr_\bfe(X)$. Since $X$ can be lifted, we can write $X_\alpha:X_i\to X_j$ as block matrix consisting of linear maps $X_{(\alpha,\chi)}:X_{(i,\chi)}\to X_{(j,\chi+e_\alpha)}$. Let $u_{(i,\chi)}\in U_{(i,\chi)}$ and $\alpha:i\to j\in Q_1$. Then we have $X_\alpha(u_{(i,\chi)})\in U_{(j,\chi+e_\alpha)}$ because $X$ can be lifted to the universal cover. Then it follows
\[X_\alpha(t.u_{(i,\chi)})=X_\alpha(t^{d(i,\chi)}u_{(i,\chi)})=t^{d(i,\chi)}X_\alpha(u_{(i,\chi)})=t^{d(i,\chi)}u_{(j,\chi+e_\alpha)}=t^{-c_\alpha}t.u_{(j,\chi+e_\alpha)}\]
for some $u_{(j,\chi+e_\alpha)}\in U_{(j,\chi+e_\alpha)}$.
Now we have \[\bigoplus_{\chi\in A_Q} t.U_{(j,\chi)}=\bigoplus_{\chi\in A_Q}t^{c(j,\chi)}t.U_{(j,\chi)}\]
for every $t\in\C^\ast$ if $c(j,\chi)$ is constant along $(j,\chi)\in \{j\}\times A_Q$. This yields  
\[t.U_j=\bigoplus_{\chi\in A_Q} t.U_{(j,\chi)}=\bigoplus_{\chi\in A_Q}t^{-c_\alpha}t.U_{(j,\chi)}\]
for every $t\in \C^\ast$.
\end{proof}

\begin{lemma}\label{degreecondition} Let $X$ be an indecomposable representation of $\hat Q$. There exists $d:\mathrm{supp} (X)\to\ZZ$ and $c_\alpha\in\mathbb N_+$ for each $\alpha\in Q_1$ such that $d(i,\chi)\ne d(j,\chi')$ if $i\neq j$ or $\chi\neq \chi'$ and such that $d(j,\chi')-d(i,\chi)=c_\alpha$ if and only if $\chi'=\chi+e_\alpha$ and $(i,\chi),\,(j,\chi+e_\alpha)\in \mathrm{supp}(X)$. 
\end{lemma}
\begin{proof}
Since $X$ is finite-dimensional, $\mathrm{supp}( X)$ is a connected finite tree. In order to prove the statement, we may assume that $d(q,0)=0$ and that $(q,0)\in \mathrm{supp}(X)$. We now fix numbers $c_\alpha$ and show that the two conditions are satisfied. To do so, let $K$ be the maximal length of a path in $\mathrm{supp}(X)$ and assume that $|Q_1|=n$. Set $c_{\alpha_1}=1$ and inductively 
\[c_{\alpha_n}> (K+1)\sum_{i=1}^{n-1}c_{\alpha_i}.\]
Note that if $A_Q$ is the free abelian group in the generators $\alpha_1,\ldots,\alpha_n$ and $f:A_Q\to \ZZ$ the group homomorphism defined by $f(\alpha)=c_\alpha$ for all $\alpha\in Q_1$,  we have $d(q,\chi)=f(\chi)$. In particular, there exist $d_i$ with $|d_i|\leq K$ such that we have
\[d(q,\chi)=\sum_{i=1}^nd_ic_{\alpha_i}.\]

First note that we obviously have $d(j,\chi+\alpha)-d(i,\chi)=c_\alpha$. Now assume that $d(i,\chi)=d(j,\chi')$ which means that $d(i,\chi)=\sum_{i=1}^{n}d_ic_{\alpha_i}=\sum_{i=1}^{n}e_ic_{\alpha_i}=d(j,\chi')$
which implies
\[\sum_{i=1}^{n-1}(d_i-e_i)c_{\alpha_i}=(e_{n}-d_{n})c_{\alpha_{n}}.\]
Since we have $|d_i-e_i|\leq K$, we obtain $$|\sum_{i=1}^{n-1}(d_i-e_i)c_{\alpha_i}|\leq K\sum_{i=1}^{n-1}c_{\alpha_i}<c_{\alpha_n}$$
This inductively yields $d_i=e_i$ for $i=n,\ldots,1$ by the choice of $c_{\alpha_{i}}$. An analogous proof applies to show that $d(j,\chi')-d(i,\chi)\neq c_\alpha$ if $\chi'\neq \chi+e_\alpha$ using that, actually, $c_{\alpha_n}> (K+1)\sum_{i=1}^{n-1}c_{\alpha_i}$.

\end{proof}
In the following, we say that $d:\mathrm{supp}(X)\to\ZZ$ satisfies the degree condition for $X$ if it has the properties of Lemma \ref{degreecondition}.
\begin{theorem}Let $X$ be a representation which can be lifted to $\hat Q$ and choose $d:\mathrm{supp}(X)\to \ZZ$ such that it satisfies the degree condition for $X$. Then the torus action of $T$ on $\bigoplus_{q\in Q_0}X_q$ induced by $t.x_{(q,\chi)}=t^{d(q,\chi)}x_{(q,\chi)}$ for $x_{(q,\chi)}\in X_{(q,\chi)}$ induces
 a torus action on $Gr_\bfe^Q(X)$ such that
  $$Gr^Q_\bfe(X)^T\cong\bigsqcup_{[\hat\bfe]} Gr^{\hat Q}_{\hat\bfe}(\hat X)$$ where $\hat\bfe$ runs through all equivalence classes of dimension vectors compatible with $\bfe$.
\end{theorem}
\begin{proof}
\end{proof}
The next step is to iterate the torus action keeping in mind the following idea: every representation, which lifts to the universal covering quiver, also lifts to the universal abelian covering quiver and also to the iterated universal abelian covering quivers, i.e. to $\hat Q^{(k)}:=\widehat{\hat Q^{(k-1)}}$ with $\hat Q=\hat Q^{(1)}$. Now it is straightforward to check that there exist natural surjective morphisms $f_k:\tilde Q\to \hat Q^{(k)}$ which become injective on finite subquivers if $k\gg 0$, see also \cite[Section 3.4]{wei}. Since the support of $X$ is finite as a representation of $\tilde Q$, we can find $k\geq 1$ such that the full subquiver with vertices $\mathrm{supp}(X)\subseteq \hat Q^{(k)}_0$ is a tree. Thus, if $\hat X^{(k)}$ is the lift of $X$ to $\hat Q^{(k)}$, there exists a torus action on the vector spaces $\hat X^{(k)}_q$ which induce a torus action on the quiver Grassmannian $Gr_{\bfe}^{\hat Q^{(k)}}(\hat X^{(k)})$. If we denote the iterated torus fixed points by $Gr^Q_\bfe(X)^{T^{(k)}}$. We obtain:
\begin{corollary}
Let $X$ be a representation which can be lifted to $\tilde Q$. Then there exists an iterated torus action such that
  $$Gr^Q_\bfe(X)^{T^{(k)}}\cong \bigsqcup_{[\hat\bfe^{(k-1)}]} Gr_{\hat\bfe^{(k-1)}}^{\hat Q^{(k-1)}}(\hat X^{(k-1)})^T\cong \bigsqcup_{[\hat\bfe^{(k)}]} Gr^{\hat Q}_{\hat\bfe^{(k)}}(\hat X^{(k)})\cong \bigsqcup_{[\tilde\bfe]} Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$$ where $\hat\bfe^{(k-1)},\,\hat\bfe^{(k)},\,\tilde\bfe$ run through all equivalence classes of dimension vectors compatible with $\bfe$.
\end{corollary}
\begin{corollary}
  affine bundles over $Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$ if $Gr^Q_\bfe(X)$ is smooth
  \[\{U\in Gr^Q_\bfe(X):\lim_{t\to0} t\cdot U\in Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)\}\]
\end{corollary}

@@

\bibitem{sch} Schofield, A.: General representations of quivers. Proceedings of the London Mathematical Society (3) \textbf{65}(1), 46-64 (1992).
\bibitem{wei} Weist, T.: Localization in quiver moduli spaces.
\end{thebibliography}
